---
title: 'Assignment 2: Movie Dataset'
author: "Mary Anna Kivenson"
date: "2/5/2019"
output:
  prettydoc::html_pretty:
    css: styles.css
    highlight: github
    theme: hpstr
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plyr)
library(reshape)
library(DT)
library(corrplot)
library(RPostgreSQL)
```


### Introduction
For this assignment, 15 people were surveyed about six recent movies. The movies included in the survey are:

* Incredibles 2
* Avengers: Infinity War
* Black Panther
* Birdbox
* A Quiet Place
* Crazy Rich Asians

The results of this survey were collected using Google Forms and participants were asked to rate each movie using the following scale:

* 5 (Excellent)
* 4 (Good)
* 3 (Okay)
* 2 (Bad)
* 1 (Terrible)

The libraries that will be used for this assignment are RpostgreSQL, ggplot2, plyr, reshape, corrplot, and dt. 



### Load SQL Library
Use the RPostgreSQL library to load the SQL Server. 
```{r libraries, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, 
                 dbname = "movies", 
                 host="localhost", 
                 user="use", 
                 password="password");
```



### Load Movies Dataframe from SQL Database
Load and query the movies database from SQL. The SQL query is located inside this github repository. The default movies dataframe is printed below.
```{r movies-dataframe}
movies_default <- dbGetQuery(con, 
                             statement = paste('SELECT * FROM movies'));
datatable(movies_default, 
     class = 'cell-border stripe', 
     options = list(
       columnDefs = list(list(className = 'dt-center', targets ="_all"))))
```


### Tranform Movies Dataframe
The format of the current dataframe is not useful for plotting (although it is ideal for correlation plots). We transform the dataframe using the reshape package - this gives each row a unique id-variable combination.
```{r movies-transform}
movies_default$personid <- (1:nrow(movies_default))
movies <-melt(movies_default, id=c("personid"))
colnames(movies) <- c("personid", "movie", "rating")
datatable(movies, 
          class = 'cell-border stripe', 
          rownames = FALSE, 
          options = list(
            columnDefs = list(
              list(className = 'dt-center', targets ="_all"))))
```


#### Plotting Ordinal Variables
The movie ratings are ordinal variables, so we can use the factor function to encode ratings as a factor. In order to plot ratings, we create another vector that only contains the numeric movie rating.
```{r movies-rating}
movies$rating <- factor(movies$rating, 
                        levels = rev(c("5 (Excellent)",
                                       "4 (Good)",
                                       "3 (Okay)",
                                       "2 (Bad)",
                                       "1 (Terrible)")))
levels(movies$rating)
movies$ratingnum <- as.numeric(substr(movies$rating, 1, 1))
```




### Movie Rating Summaries
First, we can summarize the movie ratings using mean and median centrality measures.

```{r movies-mean}
meantable <- setNames(aggregate(movies$ratingnum, 
                                list(movies$movie), mean), 
                      c("movie","mean"))
mediantable <- setNames(aggregate(movies$ratingnum, 
                                  list(movies$movie), median), 
                        c("movie","median"))
datatable(merge(x = meantable, y = mediantable, 
                by = "movie", all.x = TRUE))
```

### Visualizing Movie Ratings

**Movie Rating Counts**

The table below displays the amount of each category of ratings by movie. We can see that the perception of all the movies is generally positive or neutral. Black Panther has the highest count of excellent ratings, followed by Avengers: Infinity War.

```{r movies-bar}
ggplot(movies, aes(x=movie, fill= rating)) + 
  geom_bar() + 
  theme(legend.position="bottom") + 
  labs(title ="Count of Rating by Movie", x = "Movie", y = "Count")
```


**Rating Distributions**

This next table shows the distribution of rating categories for each movie. We can see the following movies have the highest proportion of **okay** ratings:

* A Quiet Place
* Incredibles 2
* Crazy Rich Asians

Bird Box had the highest proportion of **good** ratings, and most surveyors rated Black Panther and Avengers: Infinity War as **excellent**.

```{r movies-density}
ggplot(movies, aes(x=ratingnum, color=movie)) + 
  geom_density() + 
  labs(title ="Distribution of Ratings by Movie", 
       x = "Rating", 
       y = "Proportion") 
```


### Recommending Movies
Based on the survey responses, how do we determine whether someone will enjoy one of these movies? Using the correlation plot below, we can make recommendations using movies that a person has already watched and enjoyed.

For example:

* If someone enjoyed Avengers: Infinity War, they are likely to enjoy Black Panther
* If someone enjoyed Black Panther, they are likely to enjoy Crazy Rich Asians
* If someone enjoyed Bird Box, they may also enjoy A Quiet Place

```{r movies-cor}
movies_default$personid <- NULL
for(x in names(movies_default)){
  movies_default[[x]] <- as.numeric(substr(movies_default[[x]], 1, 1))
  }
moviescor <- cor(movies_default)
corrplot(moviescor, 
         method="color",  
         addCoef.col = "black",  
         tl.col = "black", 
         tl.srt = 90, 
         col = cm.colors(100))
```
  
  