---
title: "Magrittr: Pipe Operators"
author: "Mary Anna Kivenson"
date: "4/20/2019"
output:
  html_document:
    highlight: pygments
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidyverse: Magrittr

```{r library, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(tidyr)
library(DT)
library(data.table)
```

Magrittr is the source of pipe operators in R. Although the most commonly used pipe operator, %>% is now available in dplyr, it also has other interesting pipe options.

### %>% Operator

To begin testing Magrittr, we import a [Heart Disease dataset](https://www.kaggle.com/ronitf/heart-disease-uci/version/1) exported from Kaggle.

```{r original, message=FALSE, warning=FALSE}
heart <- read.csv("https://raw.githubusercontent.com/mkivenson/Data-Acquisition-and-Management/master/TidyVerse/heart.csv")
datatable(heart)
```

Let's say that we want to perform the following data wrangling techniques on the dataset:

* Select only the age, sex, trestbps, and chol variables.
* Filter for younger patients with an age of less than or equal to 50.
* Filter for female patients.
* Gather trestbps and chol columns into rows
* Sort by age

Using dpylr, tidyr, and data.table alone (without piping operators), this code would look something like this:

```{r nopipes, message=FALSE, warning=FALSE}
data <- select(heart, age, sex, trestbps, chol)
data <- filter(data, (age <= 50) & (sex == 0))
data <- gather(data, "metric", "value", 3:4)
data <- arrange(data, age)
datatable(data)
```

If we instead use the magrittr %>% pipe operator, we can do this. It is the same result but somewhat cleaner. 

```{r pipe, message=FALSE, warning=FALSE}
data <- heart %>% 
  select(age, sex, trestbps, chol) %>%
  filter((age <= 50) & (sex == 0)) %>%
  gather("metric", "value", 3:4) %>%
  arrange(age)

datatable(data)
```

### %<>% Operator

What if we wanted to modify the mtcars dataframe directly, rather than creating a new dataframe (or writing heart <- heart)? We can just use the %<>% operator instead of %>%:

```{r pipeassign, message=FALSE, warning=FALSE}
heart %<>% 
  select(age, sex, trestbps, chol) %>%
  filter((age <= 50) & (sex == 0)) %>%
  gather("metric", "value", 3:4) %>%
  arrange(age)

datatable(heart)
```

### Potential Expansion

Other magrittr operators include the %T>% and %$% pipes.