---
title: "Assignment 4"
author: "Mary Anna Kivenson"
date: "2/25/2019"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)
library(zoo)
```

## Importing Data
Import a .csv file that uses a "wide” structure (see below). We will perform tidying and transformations on this data.
```{r data}
flights <- read.csv("https://raw.githubusercontent.com/mkivenson/Data-Acquisition-and-Management/master/Assignment%204/airlines.csv", na.strings="")
flights
```

## Tidying and Transforming
Using the dplyr and zoo libraries, we tidy this data so that:

* Columns for different cities become one column  

```{r t1, echo=FALSE}
flights <- (gather(flights,"city","count","Los.Angeles","Phoenix","San.Diego","San.Francisco","Seattle"))
```

* All columns have column names  

```{r t2, echo=FALSE}
colnames(flights) <- c('airline','status','city','count')
```

* All rows are populated with airline information  

```{r t3, echo=FALSE}
flights <- flights %>% do(na.locf(flights))
```

* All rows are distinct  

```{r t4, echo=FALSE}
flights <- flights %>% distinct()
```

This gives us the tidy data table you see below.   

```{r t5, echo=FALSE}
datatable(flights)
```