---
title: "Lab 2.3"
author: "Mary Anna Kivenson"
date: "2/5/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load SQL Library
Install the RPostgreSQL library and load the SQL Server.
```{r libraries}
library(RPostgreSQL);
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "tb", host="localhost", user="user", password="password");
```


## Load Population Dataframe
Load the population dataframe from github.
```{r population-dataframe}
population <- read.csv('https://raw.githubusercontent.com/mkivenson/Data-Acquisition-and-Management/master/Lab%202.3/population.csv', header=TRUE, sep=',')
head(population)
```

## Load tb Dataframe from SQL Database
Load and query the tb database from SQL.
```{r tb-dataframe}
tb <- dbGetQuery(con, statement = paste(
  'SELECT country, year, SUM(child + adult + elderly) AS cases',
  'FROM tb',
  'WHERE child + adult + elderly != 0',
  'GROUP BY country, year',
  'ORDER BY country, year;'));
head(tb, n=3)
```

## Left join tb and population databases
Left join the tb and population databases by country and year.
```{r left-join}
join_tb<- merge(x = tb, y = population, by = c("country", "year"), all.x = TRUE)
head(join_tb, n=3)
```

## Add column calculating rate of tb occurance
Calculate the rate of tb occurance by dividing the amount of tb cases by total population.
```{r rate-tb}
join_tb$ratetb <- join_tb[,3]/join_tb[,4]
```

##tb occurance summary
Summaries by country of tb incidence occurance and rate. 

**Countries with highest mean tb incidence rates**
```{r summary-tb-1}
tb_by_country_top <- sort(tapply(join_tb$cases, join_tb$country, mean),decreasing=TRUE)
head(tb_by_country_top)
```

**Countries with highest mean tb occurance per population**
```{r summary-tb-2}
tbrate_by_country_top <- sort(tapply(join_tb$ratetb, join_tb$country, mean),decreasing=TRUE)
head(tbrate_by_country_top)
```

**Annual mean tb incidence rates**
```{r summary-tb-3}
barplot(tapply(join_tb$cases, join_tb$year, mean))

```

**Annual mean tb occurance per population**
```{r summary-tb-4}
barplot(tapply(join_tb$ratetb, join_tb$year, mean))

```
