---
title: "New York Times Web APIs"
author: "Mary Anna Kivenson"
date: "3/30/2019"
output:
  html_document:
    highlight: pygments
    theme: united
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
In this exercise, JSON data from New York Times APIs are read and transformed into an R dataframe. 

## Books API
The New York Times provides an API of its best sellings books, which is the first dataset we use. 
```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(httr) #for httr requests
library(jsonlite)
library(tidyverse)
library(DT)
library(ggplot2)
library(magrittr)
```

### Request
First, we request data on current NYT Best Sellers. We use the httr library to retrieve the response specified by the API url. 
```{r request, message=FALSE, warning=FALSE}
url <- "https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json?api-key=f5RHWlYOgGenjKxrLrkIkWLfVGjbZM9D"
response <- GET(url) # Retrieve whatever is specified by the URL 
response # Server response returned by function
```

### Content
We then use the jsonlite package to transform the books json file into a dataframe. There is a nested json column, "isbns", that contains various types of isbns. We expand that column and then select relevent columns to preview the data. 
```{r response, message=FALSE, warning=FALSE}
json <- content(response, as="text")
res <- jsonlite::fromJSON(json)
books <- res[["results"]][["books"]]
colnames(books)

books$isbn10 <- books %>% select("isbns") %>% apply(1, FUN=function(v) v[["isbns"]][["isbn10"]])
books$isbn13 <- books %>% select("isbns") %>% apply(1, FUN=function(v) v[["isbns"]][["isbn13"]])

books %<>%
unnest(isbn10, isbn13) %>% select(c("rank", "weeks_on_list", "author", "title", "amazon_product_url", "isbn10", "isbn13"))
datatable(books)
```

## Popular Articles API
There is another API that contains data on most popular New York Times articles in terms of amount of times emailed, shared on Facebook, and viewed.

### Most Emailed
```{r Emailed, message=FALSE, warning=FALSE}
emailed <- content(GET("https://api.nytimes.com/svc/mostpopular/v2/emailed/7.json?api-key=f5RHWlYOgGenjKxrLrkIkWLfVGjbZM9D"), as="text")
emailed <- jsonlite::fromJSON(emailed)
datatable(as.data.frame(emailed))
```


### Most Shared
```{r Shared, message=FALSE, warning=FALSE}
shared <- content(GET("https://api.nytimes.com/svc/mostpopular/v2/shared/1/facebook.json?api-key=f5RHWlYOgGenjKxrLrkIkWLfVGjbZM9D"), as="text")
shared <- jsonlite::fromJSON(shared)
datatable(as.data.frame(shared))
```

### Most Viewed
```{r Viewed, message=FALSE, warning=FALSE}
viewed <- content(GET("https://api.nytimes.com/svc/mostpopular/v2/viewed/1.json?api-key=f5RHWlYOgGenjKxrLrkIkWLfVGjbZM9D"), as="text")
viewed <- jsonlite::fromJSON(viewed)
datatable(as.data.frame(viewed))
```